\documentclass[10pt,twoside,english,a4paper]{article}
\usepackage[english]{babel}
%\usepackage[T1]{fontenc}
\usepackage[IL2]{fontenc} % lepšia sadzba písmena Ľ než v T1
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{url} % príkaz \url na formátovanie URL
\usepackage{hyperref} % odkazy v texte budú aktívne (pri niektorých triedach dokumentov spôsobuje posun textu)

\usepackage{cite}
%\usepackage{times}

\pagestyle{headings}

\title{Algorithms for improving image processing in mobile photography \thanks{Semestrálny projekt v predmete Metódy inžinierskej práce, ak. rok 2021, vedenie: Artem Vasko}} % meno a priezvisko vyučujúceho na cvičeniach

\author{Artem Vasko\\[2pt]
	{\small Slovenská technická univerzita v Bratislave}\\
	{\small Fakulta informatiky a informačných technológií}\\
	{\small \texttt{xvaskoa1@stuba.sk}}
	}

\date{\small 9. september 2021} % upravte



\begin{document}

\maketitle

% \begin{abstract}

% \end{abstract}



\section{Introduction}

Today slightly every smartphone has a camera. It is very convenient,
% may be percentage of all photos made on phones vs cameras
until we speak of quality of photos, captured with phone.
By this article I will suggest way to improve quality by some algorithms and
methods for post-processing that can be included to camera
app. I choose this topic because as photograph myself, I was always
unpleasant by my photos taken by phone. It maybe arranges for some people,
but after seeing photos taken on beginner-class camera I struggled on
insufficient quality of taken images. So, this article would add some
possible solutions to the problem of poor dynamic range of photos
taken with mobile devices nowadays.



\section{What are possible solutions?} \label{nejaka}

How can we solve problem with bad quality of photos? There are two ways:
by improving hardware or improving software. But why just don't install
large sensors (with big pixels in it) from the cameras to the smartphones,
making a huge hardware improvement?
\newline It all because of dimensions of the smartphones. Today it's no
way that big sensors can be handled in smartphone, which width is mostly
limited to  10mm. If you know how camera work, you maybe recently
understand the reason why it so, and it is not the size of sensor, but the
size of the lens. The sensor is not a solar panel-like device whose size
determines how much light you collect. The sensor is behind the lens
and in fact, the lens is the determining factor.
A large diameter lens captures a lot of light.
It almost doesn't matter how large the sensor is. And by installing large
sensor at once we should install lens with huge focal length. This is the
main issue for relatively thin smartphone.
\newline *img of camera structure*
\newline In simple words, larger sensor (with big pixels in it) will require
larger optics in terms of aperture and focal length.
This means that the body of the phone will need to be thicker in
order to provide the required distance between the lens and the sensor.
\newline
% To solve this problem, we need to know the specific reason.
So if we can not just increase the sensors size, we need to go in another
way, namely improve software. The biggest problem is lack of tones and
colors, or in other words, lack  of dynamic range. Especially in
high-contrast photos.
To understand what dynamic range is, we can imagine it as a piano keyboard. There is
a range of tones, in which we can see like piano has keyboard with limited
width, through which we can play. And the wider
it is, the better. Even human eye can not see all range at once, but sensors
with big pixels in
digital cameras can because of manual exposure, so we can put exposure
exactly where we want it to be. Only the biggest and most expensive
sensors can accommodate all 12 stops (there is a 12-stop difference in the
highlights and shadows). But we are talking about matrices with small
pixels, by which we want to cover all range of tones.
\newline The simplest decision is to combine few under- and over-exposed
photo into one frame. Doing that we create HDR photo. \cite{HDR1}



% \newline Z obr.~\ref{f:rozhod} je všetko jasné. 

% \begin{figure*}[tbh]
% 	\caption{Combining few photos with different exposures 
% 	usually creates HDR image}
% 	\label{f:rozhod}
% \end{figure*}



\section{HDR photos} \label{ina}
Yes, HDR photographs can solve the problem and
make dynamic range wider. But like every technology
it has some problems.

Main problem is difference between over- and under-exposed images...
% describe fuul problem?
Firstly lets see some explanation (expl.~\ref{ina:nejake}),
and then one more (expl.~\ref{ina:este}) to fully understand
why is it happening.
\newline With that knowledge in head we can conclude that  main problem is
difference in images, captured in a)a different time, b)with different
important settings, such as shutter-speed.
\footnote{The speed at which
	the shutter of the camera closes.}

% Môže sa zdať, že problém vlastne nejestvuje\cite{Coplien:MPD}, ale bolo dokázané, že to tak nie je~\cite{Czarnecki:Staged, Czarnecki:Progress}. Napriek tomu, aj dnes na webe narazíme na všelijaké pochybné názory\cite{PLP-Framework}. Dôležité veci možno \emph{zdôrazniť kurzívou}.


\subsection{Explanation} \label{ina:nejake}

Some fact to remember:
\begin{itemize}
	\item Dynamic range is the range between the maximum and minimum
	      measurable light intensities.
	\item Combining few photos with different exposures
	      usually creates HDR image
	      %   \begin{itemize}

	      %   \end{itemize}
\end{itemize}


% \begin{enumerate}
% 	\item jedna vec
% 	\item druhá vec
% 	      \begin{enumerate}
% 		      \item x
% 		      \item y
% 	      \end{enumerate}
% \end{enumerate}


\subsection{Explanation} \label{ina:este}

\paragraph{How are made under/over-exposed images}
\begin{itemize}
	\item Over- and under-exposed images are made by decreasing
	      or respectively increasing shutter-speed in camera.
\end{itemize}



\section{Under-exposing images} \label{dolezita}

Problems start with moving objects in frame. Different shutter-speed causes
the difference in frames that in turn combines into final picture. 
Also, over-exposed images has more distortion because shutter-speed 
is relatively slow.
To remove the difference Google \cite{Google}
scientists
firstly changed few photos with different exposure to few underexposed ones.
That helped to achieve more similar primary photos, remove photos that 
captured on slow shutter-speed and in this way to better merge them into one.
% more detailed later
\newline ...But there is one more problem: several frames even on the same exposure
still need some time to make. In motion, it becomes mostly impossible.



\section{Burst photographs and buffer} \label{dolezitejsia}
Another possible decision from Google Research is buffer. So, at once
you open the camera app, it starts to make burst photograph. \cite{Burst}
It keeps doing it all the time, before someone make a photo. 
By this way, some last photos from buffer are compare with the photo made on 
the moment the shutter was pressed. Then from all photographs picking the 
sharpest image. This approach is commonly known as 
lucky imaging  \cite{LuckySharp}. 
\newline Also, in modern smartphones there are usually more than one camera, 
and my suggestion is to improve the number of burst photographs by 
at the time of creating a shot, take images from three modules at once.

\section{Conclusions} \label{zaver} % prípadne iný variant názvu
I described a system for improving quality of photo by 
taking few underexposed frames from buffer and then to combine them all to 
one image. It possibly can increase dynamic range of photo and also 
to produce better resolution. Also, I counted disadvantages of modern HDR 
technology and offered a replacement. 	

\cite{HDR2}
%\acknowledgement{Ak niekomu chcete poďakovať\ldots}

% týmto sa generuje zoznam literatúry z obsahu súboru literatura.bib podľa toho, na čo sa v článku odkazujete




\bibliography{literatura}
\bibliographystyle{plain} % prípadne alpha, abbrv alebo hociktorý iný
\end{document}